---
title: "Getting Started"
output: 
  md_document:
    variant: gfm
---

```{r}
library(tidyverse)
```

It is assumed that users of this data may have different levels of proficiency with different data tools.  Below are a few quick notes on how to access the data and start using it.

# Cloning the repo locally

If you cloned the GitHub repository on your local machine and have R installed, you can simply run the following:

```{r load, eval=FALSE}
tmi <- read_csv("data/tmi.csv")
# atu <- read_csv("data/atu.csv")
aat <- read_csv("data/aat.csv")
```

# Pulling from GitHub

If you want to pull the raw flat files from the web using another coding language (e.g. Python), use the following steps:

1. Go to the [data folder of this repo](https://github.com/j-hagedorn/trilogy/tree/master/data)
2. Select the data file you want to access (e.g. `tmi.csv`)
3. Click the "View raw" button to display the raw data format.
4. Copy the URL from the browser.

You can then read the data into R, Python, or other tools.

# Downloading to Excel

Users who are comfortable using MS Excel can access the raw flat files using the URLs mentioned above and [import them into Excel](https://support.office.com/en-za/article/Import-or-export-text-txt-or-csv-files-5250ac4c-663c-47ce-937b-339e391393ba) using the steps outlined in the link.
