---
title: "Notes for AFT Article"
output: html_document
date:  '`r Sys.Date()`'
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = F, warning = F, message = F)
aft <- read_csv("../../data/aft.csv")
tmi <- read_csv("../../data/tmi.csv")
atu <- read_csv("../../data/atu.csv")
```

# Cleaning

-   Clean up tale names regex in `atu`
-   Finish ATU sequences
-   Clean up ID matches between `atu` and `aft`
-   Reconcile use of older ATU ID by some `aft` texts

```{r}

aft <- 
  aft %>%
  mutate(
    atu_id = str_remove(atu_id, "^0+"),
    atu_id = str_to_upper(atu_id),
    atu_id = recode(
      atu_id,
      `676`='954',
      `333A`='333',
      `6070B`='113A',
      `47E`='47B',
      `285D`='285A',
      `170A`='2034F'
    )
  ) 

```


# Exploratory analysis: What's in the AFT corpus?

## Tale Types

-   Proportion of ATU represented by `aft`

```{r unmatched_in_atu, eval=FALSE, include=FALSE}

tst <-
  aft %>%
  select(atu_id,type_name) %>%
  mutate(
    atu_id = str_remove(atu_id, "^0+"),
    atu_id = str_to_upper(atu_id)
  ) %>%
  anti_join(atu %>% select(chapter:tale_name), by = "atu_id") %>%
  distinct()

```

```{r merge_aft_atu}

# Merge AFT with ATU and filter unmatched items from AFT

aft_in_atu <-
  aft %>%
  select(atu_id,type_name) %>%
  mutate(in_aft = T) %>%
  full_join(atu %>% select(chapter:tale_name), by = "atu_id") %>%
  filter(!is.na(chapter)) %>% # remove items in AFT but not ATU
  mutate(in_aft = ifelse(!is.na(in_aft),T,F)) %>%
  group_by(atu_id,chapter,division,sub_division,tale_name) %>%
  summarize(n_tales = sum(in_aft))

```

### By ATU Chapter/Division

Summary stats by ATU chapter:

```{r}
library(formattable)

aft_in_atu %>%
  mutate(with_tale = n_tales > 0) %>%
  group_by(chapter) %>%
  summarize(
    n_types = n_distinct(atu_id),
    n_types_with_tale = sum(with_tale),
    n_tales = sum(n_tales)
  ) %>%
  mutate(
    pct_with_tales = round(n_types_with_tale / n_types * 100, digits = 1),
    tales_per = round(n_tales / n_types_with_tale, digits = 1)
  ) %>%
  arrange(desc(tales_per)) %>%
  select(-n_types_with_tale) %>%
  formattable(
    list(
      n_tales = color_bar("#1BB6AFFF"),
      # n_types_with_tale = color_bar("#EC921DFF"),
      n_types = color_bar("#1BB6AFFF"),
      pct_with_tales = color_tile("transparent", "#EF562AFF"),
      tales_per = color_tile("transparent", "#EF562AFF")
    )
  )


```

The treemap below shows the nested sets of the ATU into which AFT texts fall, by `chapter`, `division`, and `sub_division`.

```{r}

aft_tree <-
  aft_in_atu %>%
  group_by(chapter,division) %>%
  summarize(
    n_types = n_distinct(atu_id),
    n_tales = sum(n_tales)
  ) %>%
  filter(!is.na(division)) %>%
  rename(parent = chapter, labels = division) %>%
  bind_rows(
    aft_in_atu %>%
    filter(!is.na(division)) %>%
    group_by(chapter) %>%
    summarize(
      n_types = n_distinct(atu_id),
      n_tales = sum(n_tales)
    ) %>%
    mutate(parent = "root") %>%
    rename(labels = chapter)
  ) %>%
  bind_rows(
    aft_in_atu %>%
    filter(!is.na(division)) %>%
    group_by(division, sub_division) %>%
    summarize(
      n_types = n_distinct(atu_id),
      n_tales = sum(n_tales)
    ) %>%
    rename(parent = division, labels = sub_division)
  ) # %>%
  # bind_rows(
  #   aft_in_atu %>%
  #   filter(!is.na(division)) %>%
  #   group_by(sub_division, tale_name) %>%
  #   summarize(
  #     n_types = n_distinct(atu_id),
  #     n_tales = sum(n_tales)
  #   ) %>%
  #   rename(parent = sub_division, labels = tale_name)
  # )

plotly::plot_ly(
  aft_tree,
  type = 'treemap',
  labels = ~labels,
  parents = ~parent,
  values = ~n_tales
)

```

## Content

### Entities

- Find a well performing model for NER (cf. https://huggingface.co/flair/ner-english-ontonotes-fast, https://huggingface.co/stanfordnlp/stanza-en)
- Find a way to label the name of a main character (example Miller, prince, etc.) as more than simply a noun
- Pull out subject and predicate from sentence.  Explore verbs related to subjects.

### Common phrases 


```{python}

```


- TextRank
- collocation/word frequency

### Topic modeling

- Define cleaning tasks and stop words to improve topic models performance; right now they are too close together, with a few main clusters of topics that are difficult to distinguish

